---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##Loading libraries required for analysis
```{r}
library(dplyr)
```

## Loading and preprocessing the data
Creating a data directory if it does not exist
```{r}
  if(!file.exists("Data")){
    dir.create("Data")
  }
```

Unzipping source file
```{r}
  unzip('activity.zip',exdir='./Data/')
```

Importing source file
```{r}
  Import <- read.csv(file='./Data/activity.csv',colClasses = c("numeric","Date","numeric") )
```

## What is mean total number of steps taken per day?

```{r}
Steps_Total <- Import %>%
                filter(steps != 'NA') %>%
                group_by(date) %>%
                summarise(Steps_Total = sum(steps))

Steps_Mean <- Steps_Total %>%
                summarise(
                          Steps_Mean = mean(Steps_Total)
                          ,Steps_Median = median(Steps_Total)
                          )
```
Creating a historgram of Total step by day
```{r}
png(filename = "./figure/Histogram_TotalStepsByDay.png")
hist(Steps_Total$Steps,xlab= "Total Steps by Day",main = "Histogram of Total Steps by Day")
dev.off()
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
