---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##Loading libraries required for analysis
```{r}
library(dplyr)  #for easy data manipulation
library(lattice) # for easy graphing
```

## Loading and preprocessing the data
Creating a data directory if it does not exist
```{r}
  if(!file.exists("Data")){
    dir.create("Data")
  }
```

Unzipping source file
```{r}
  unzip('activity.zip',exdir='./Data/')
```

Importing source file
```{r}
  Import <- read.csv(file='./Data/activity.csv',colClasses = c("numeric","Date","numeric") )
```

## What is mean total number of steps taken per day?

```{r}
Steps_Total <- Import %>%
                filter(steps != 'NA') %>%
                group_by(date) %>%
                summarise(Steps_Total = sum(steps))

Steps_Mean <- Steps_Total %>%
                summarise(
                          Steps_Mean = mean(Steps_Total)
                          ,Steps_Median = median(Steps_Total)
                          )

Step_Mean <- round(as.numeric(Steps_Mean[1,"Steps_Mean"]),1)
Step_Median <- round(as.numeric(Steps_Mean[1,"Steps_Median"]),1)


```
Creating a historgram of Total step by day
```{r}
hist(Steps_Total$Steps,xlab= "Total Steps by Day",main = "Histogram of Total Steps by Day")
```
  
The following figures are calculated from the total number of steps per day  
* mean: `r Step_Mean`  
* median: `r Step_Median`  
  
## What is the average daily activity pattern?
```{r}
Steps_Time <- Import %>%
                filter(steps != "NA") %>%
                group_by(interval) %>%
                summarise(Steps_Mean = mean(steps))

plot(Steps_Time$Steps_Mean~Steps_Time$interval,type="l",xlab="interval",ylab="Mean Steps")

Steps_Time_Max <- Steps_Time[Steps_Time$Steps_Mean == max(Steps_Time[,2]),]

```
  
The maximum number of steps, averaged across all days, is taken during `r Steps_Time_Max[1,"interval"]` interval.  
  
## Imputing missing values

```{r}
#calculating the number of missing rows
Missing_Steps <- nrow(Import[complete.cases(Import) != TRUE,])

#calculating average steps per interval
Steps_Interval <- Import %>%
                          filter(steps != "NA") %>%
                          group_by(interval) %>%
                          summarise(Interval_Mean = mean(steps))

#merging average steps per interval onto intial import
Missing_Prep <- merge(Import,Steps_Interval,by.x="interval",by.y="interval",all.x=TRUE)

#filling in missing steps with interval
Missing_Fixed <- mutate(Missing_Prep,steps = ifelse(is.na(Missing_Prep$steps),Missing_Prep$Interval_Mean,Missing_Prep$steps))

#calculating total steps per day in fixed dataframe
Missing_Fixed_DateTotal <- Missing_Fixed %>%
                            group_by(date) %>%
                            summarise(Date_Total = sum(steps)
                                      )

#caculating mean and median from fixed dataset
Missing_Fixed_Steps_Mean <- Missing_Fixed_DateTotal %>%
                                summarise(
                                          Steps_Mean = mean(Date_Total)
                                          ,Steps_Median = median(Date_Total)
                                          )

#generating histogram of total steps by day (fixed dataframe)
hist(Missing_Fixed_DateTotal$Date_Total,xlab="Total Steps per Day",main="Histogram Total Steps per Day (with NA's imputed)")

```
  
There are `r Missing_Steps` number of rows with missing values.  
The following figures were calculated with missing values imputed:  
* mean: `r Missing_Fixed_Steps_Mean[1,"Steps_Mean"]` (originally: `r Step_Mean`)  
* median: `r Missing_Fixed_Steps_Mean[1,"Steps_Median"]` (originally: `r Step_Median`)
  
## Are there differences in activity patterns between weekdays and weekends?
```{r}
#adding day of the week to dataframe
Missing_Fixed$Day <- weekdays(Missing_Fixed$date)

#adding whether it is a weekday or weekend
Missing_Fixed$Weekday <- ifelse(Missing_Fixed$Day == "Saturday" | Missing_Fixed$Day == "Saturday","weekend","weekday")

#grouping weekend and weekday steps
Weekday_Steps <- Missing_Fixed %>%
                  group_by(Weekday,interval) %>%
                  summarise(Steps_Interval = sum(steps))

#plotting weekend and weekday steps
xyplot(Steps_Interval ~ interval |Weekday,data=Weekday_Steps,type="l",xlab="Interval",ylab="Number of steps",layout=c(1,2))

```